---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it.

```{r}
# Libraries 
library(pacman)
p_load(tidyverse,dplyr,ggplot2,readr,metafor,lme4, lmerTest,cran)

# working directory and data 
setwd("C:/Users/Bruger/Google Drev/Cognitive Science/Experimental methods 3/assignment 5")

Data <- read.csv("C:/Users/Bruger/Google Drev/Cognitive Science/Experimental methods 3/assignment 5/Data.csv", sep=";")
View(Data)

```

```{r}

## META ANALYSIS
sum(Data$SampleSizeSchizo)
# 518
sum(Data$SampleSizeContros, na.rm = T)
# 216


# Mean distribution: calculate effect size (yi) and variance (vi) -> look up
Escalc1 <- escalc("SMD", n1i = Data$SampleSizeContros, n2i = Data$SampleSizeSchizo, m1i = Data$PitchMeanControls, m2i = Data$PitchMeanSchizo, sd1i = Data$PitchMeanControlsSD, sd2i = Data$PitchMeanSchizoSD, data = Data)

# Mean distribution
#model lmer
m1 = lmer(yi ~ 1 +(1|Article),weights = 1/vi,Escalc1,control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1)
#estimate = -0.17 
#stEr = 0.17


#model rma
m2 = rma(yi,vi,data = Escalc1, slab = Article)
summary(m2)
#tau^2 (estimated amount of total heterogeneity): 0.0607
#I^2 (total heterogeneity / total variability):   44.28%
# estimate = -0.13 
#std error = 0.19 non significant 

# forest plot
forest(m2)

# SD distribution: calculate effect size (yi) and variance (vi) -> look up
Escalc2 <- escalc("SMD", n1i = Data$SampleSizeContros, n2i = Data$SampleSizeSchizo, m1i = Data$PitchSDControls, m2i = Data$PitchSDSchizo, sd1i = Data$PitchSDControlsSD, sd2i = Data$PitchSDSchizoSD, data = Data)

# SD distribution: 
# model lmer
m3 = lmer(yi ~ 1 +(1|Article),weights = 1/vi,Escalc2,control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m3)
# estimate = 0.125
# std.error = 0.579

m4 = rma(yi,vi,data = Escalc2, slab = Article)
summary(m4)
#tau^2 (estimated amount of total heterogeneity): 4.2287 (SE = 2.7324)
#I^2 (total heterogeneity / total variability):   98.15%
#estimate = -0.35
#se = 0.85 non significant 
forest(m4)


```

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

```{r}
## PREPARE DATA 

A3_data <- read.delim("final_rqa.csv",sep = ",")

# subset control 
control_data = subset(A3_data, A3_data$diagnosis=="control")
# find sample size of control 
SampleSizeContros = length(unique(control_data$participant))
# control = 59

# subset schizo
schizo_data = subset(A3_data,A3_data$diagnosis=="schizophrenia")
# find sample size schizo
SampleSizeSchizo = length(unique(schizo_data$participant))
# schizo = 57

# mean pitch of control
PitchMeanControls = mean(control_data$mean)
# sd of mean pitch contorl
PitchMeanControlsSD = sd(control_data$mean)

# mean pitch of schizo
PitchMeanSchizo = mean(schizo_data$mean)
# sd of mean pitch schizo
PitchMeanSchizoSD = sd(schizo_data$mean)

PitchSDControls = mean(control_data$stdDev)
# mean control sd = 25.6 
PitchSDControlsSD = sd(control_data$stdDev)
# sd control sd = 17.9 # how spread the distribution of sd are. how close are the patients 

PitchSDSchizo = mean(schizo_data$stdDev)
# mean schizo sd = 24.03
PitchSDSchizoSD = sd(schizo_data$stdDev)
# sd schizo sd = 36.8

Year = 2017

Article = "A3"

A3_sum = data.frame(Article,Year,SampleSizeContros, SampleSizeSchizo, PitchMeanControls,PitchMeanControlsSD,PitchMeanSchizo,PitchMeanSchizoSD,PitchSDControls,PitchSDControlsSD,PitchSDSchizo,PitchSDSchizoSD)

final_df <- rbind(Data,A3_sum) 


```

```{r}
## CALCULATE EFFECT SIZES AND FOREST PLOT: META ANALYSIS 


# Mean distribution: calculate effect size (yi) and variance (vi) -> look up
Escalc1_final <- escalc("SMD", n1i = final_df$SampleSizeContros, n2i = final_df$SampleSizeSchizo, m1i = final_df$PitchMeanControls, m2i = final_df$PitchMeanSchizo, sd1i = final_df$PitchMeanControlsSD, sd2i = final_df$PitchMeanSchizoSD, data = final_df)

# Mean distribution
#model lmer
m5 = lmer(yi ~ 1 +(1|Article),weights = 1/vi,Escalc1_final,control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m5)
# estimate = -0.194
# std error = 0.132 

m6 = rma(yi,vi,data = Escalc1_final, slab = Article)
summary(m6)
#tau^2 (estimated amount of total heterogeneity): 0.0330 (SE = 0.0677)
#I^2 (total heterogeneity / total variability):   34.51%
#estimate = -0.16
#se = 0.14 non significant 

# forest plot
forest(m6)

# SD distribution: calculate effect size (yi) and variance (vi) -> look up
Escalc2_final <- escalc("SMD", n1i = SampleSizeContros, n2i = SampleSizeSchizo, m1i = PitchSDControls, m2i = PitchSDSchizo, sd1i = PitchSDControlsSD, sd2i = PitchSDSchizoSD, data = final_df)

# SD distribution:
# model lmer 
m7 <- lmer(yi ~ 1 +(1|Article),weights = 1/vi,Escalc2_final,control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m7)
# estimate = 0.11
# std error = 0.45 

m8 = rma(yi,vi,data = Escalc2_final, slab = Article)
summary(m8)
#tau^2 (estimated amount of total heterogeneity): 3.4938 (SE = 2.0651)
#I^2 (total heterogeneity / total variability):   98.08%
#estimate = -0.29
#se=0.71

#forest plot
forest(m8)



```


3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.
```{r}

# Heterogenity 

#mean
summary(m6)
# overall variance measure
#tau^2 (estimated amount of total heterogeneity): 0.0330 (SE = 0.0677)
# variance caused not by whithin study variation. procentage of participants which do not fall within CI of the other studies. 
#I^2 (total heterogeneity / total variability):   34.51%
# ratio of observed to whithin study variation
# Q(df = 6) = 129.5711, p-val < .0001

#sd
summary(m8)
# tau^2 (estimated amount of total heterogeneity): 3.4938 (SE = 2.0651)
# I^2 (total heterogeneity / total variability):   98.08%
  # -> the studie measures are to different, not comparable. 

# source of heterogentity: influential studies
#MEAN
inf_m6 <- influence(m6)
print(inf_m6)
plot(inf_m6)
# influential study: Martinez et al 2015

# SD
inf_m8 <- influence(m8)
print(inf_m8)
plot(inf_m8)
# influential study: Cohen et al. 2014 

# publication bias 

# funnelplot MEAN
m6_funnel <- funnel(m6, main = "Random-Effects Model",xlab = "Standardized Mean Difference") 

# rank correlation test MEAN
ranktest(m6) 
# Kendall's tau = 0.2000, p = 0.8167
# -> symetrical variance

# funnelplot SD
m8_funnel <- funnel(m8, main = "Random-Effects Model",xlab = "Standardized Mean Difference")

# rank correlation test SD
ranktest(m8)
# Kendall's tau = 0.0476, p = 1.0000
# -> symetrical variance!

```

## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=0
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2
